---
title: "Lanzarote Soil"
author: "Yago Sánchez Gancedo"
date:
output: html_document
---


## 1. Introduction of the data

The first task is to read the data in R

```{r}
install.packages("openxlsx", repos = "http://cran.us.r-project.org")
library(openxlsx)

#Creating the local path to the data, and using that path to generate an object containing such data

DataPath <- "/Users/cacharra2.0/Desktop/TFM/Lanzarotesoil.xlsx"
SoilData <- read.xlsx(DataPath, colNames = TRUE)
```

## 2. Superficial exploration of the data

Once the data is stored in the variable "SoilData", a brief superficial exploration can be performed to ensure that everything is correct and that the data is treated accordingly:

Visualization of the first and last subset of data, to check if it matches with the original data table.
```{r}
head(SoilData)
tail(SoilData)
```

Structure of the data.

```{r}
dim(SoilData)
ncol(SoilData)
nrow(SoilData)
str(SoilData)
```

A summary to see if the variables are treated with the correct format

```{r}
summary(SoilData)
```

As it can be seen with the two previous commands, some variables are not treated as desired. A correction is needed:

```{r}
SoilData$SAMPLE.NUMBER <- as.character(SoilData$SAMPLE.NUMBER)
SoilData$CLASS <- factor(SoilData$CLASS)
SoilData$SUBCLASS1_FARMED <- factor(SoilData$SUBCLASS1_FARMED)
SoilData$SUBCLASS2_MULCHING <- factor(SoilData$SUBCLASS2_MULCHING)
SoilData$SUBCLASS3_IRRIGATION <- factor(SoilData$SUBCLASS3_IRRIGATION)
SoilData$CLASS.REFERENCE <- factor(SoilData$CLASS.REFERENCE)
SoilData$DEPTH <- factor(SoilData$DEPTH) #LABELS!
SoilData$TEXTURE <- factor(SoilData$TEXTURE) #LABELS!
```

## 3. Wrangling of the data

Creating a new factor with the different pH types

```{r}
SoilData$PHtype <- cut(SoilData$PH, breaks=c(-Inf, 6, 8, Inf), labels=c("Acid", "Neutral", "Basic"))

#More Wrangling?

```

## 4. Variable exploration