---
title: "Lanzarote Soil - Functions"
author: "Yago Sánchez Gancedo"
date:
output: html_document
---


## map_plot

Generates a plot considering the geographic distribution of a variable and a subclass clasification. The format of the data points is different deppending on the nature of the variable.

```{r}
map_plot <- function(subclass_chr, subclass_label, dataframe, variable_chr){
                
                require(ggplot2)
                require(dplyr)
   
                filtered_data <- dplyr::filter(dataframe, !is.na(dataframe[,variable_chr]))
                
                if (is.numeric(dataframe[,variable_chr])){
                        
                    ggplot(filtered_data, aes(x=XUTM, y=YUTM, size=filtered_data[,variable_chr], col = filtered_data[,subclass_chr]))+
                  geom_point()+
                  ggtitle(paste(variable_chr, "according to location and", subclass_label))+
                  labs(size=variable_chr, col=subclass_label)+
                  theme(legend.title=element_text(size=9),
                        legend.text=element_text(size=7),
                        legend.position = c(0.85,0.2),
                        legend.box="horizontal",
                        legend.key.size=unit(0.05, "cm"))  
                        
                }else if(is.factor(dataframe[,variable_chr])) {
                  ggplot(filtered_data, aes(x=XUTM, y=YUTM, shape=filtered_data[,variable_chr], col = filtered_data[,subclass_chr]))+
                    geom_point()+
                    ggtitle(paste(variable_chr, "according to location and", subclass_label))+
                    labs(shape=variable_chr, col=subclass_label)+
                    theme(legend.title=element_text(size=9),
                          legend.text=element_text(size=7),
                          legend.position = c(0.85,0.2),
                          legend.box="horizontal",
                          legend.key.size=unit(0.05, "cm"))
                }
                 
}
```

## soil_boxplot

```{r}

soil_boxplot <- function(subclass_chr, subclass_label, dataframe, variable_chr){
  
  require(ggplot2)
  require(dplyr)
  
  filtered_data <- dplyr::filter(dataframe, !is.na(dataframe[,variable_chr]))
  
  ggplot(filtered_data, aes(x=filtered_data[,subclass_chr], y=filtered_data[,variable_chr]))+
    geom_boxplot()+
    ggtitle(paste("Boxplot of", variable_chr, "related to", subclass_label))+
    labs(x=subclass_label, y=variable_chr)
        
  
}



```

## variable_exploration

Given a variable, the function will generate a pdf with several outputs including statistics, plots, etc. with the aim of making a general exploration of the variable. It also takes into account the nature of the variable.


```{r}
variable_exploration <- function (dataframe, variable_chr) {
        
        require(ggplot2)
        require(dplyr)
        require(gridExtra)
        
        # the name of the resulting pdf file.
         filename <- paste("Exploration of ", variable_chr,".pdf", sep="") 
        
        # if the variable is numeric, a summary, histogram, test for normality and several "map plots" will be obtained.
     if(is.numeric(dataframe[,variable_chr])){
             
         pdf(filename)
         
         grid.table(capture.output(summary(dataframe[,variable_chr])))
         
         print(ggplot(dataframe, aes(dataframe[,variable_chr])) + geom_histogram() + ggtitle(paste("Histogram of", variable_chr)) + labs(x=(variable_chr)))
         
         plot.new()
         
         grid.table(capture.output(shapiro.test(dataframe[,variable_chr])))
         
         #map_plot("CLASS", "Soil Class", dataframe, variable_chr)??
         
         print(map_plot("SUBCLASS1_FARMED", "Soil Farming", dataframe, variable_chr))
         
         print(map_plot("SUBCLASS2_MULCHING", "Soil mulching",dataframe, variable_chr))
         
         print(map_plot("SUBCLASS3_IRRIGATION", "Soil Irrigation",dataframe, variable_chr))
         
         print(soil_boxplot("SUBCLASS1_FARMED", "Soil Farming", dataframe, variable_chr))
         
         print(soil_boxplot("SUBCLASS2_MULCHING", "Soil mulching",dataframe, variable_chr))
         
         print(soil_boxplot("SUBCLASS3_IRRIGATION", "Soil Irrigation",dataframe, variable_chr))
         
         dev.off()
         return()
     }
        
         #On the other hand, if the variable is a factor, a summary, bar plot and several "map plots" will be obtained.
         
     else if(is.factor(dataframe[,variable_chr])){
             
        pdf(filename)
             
        grid.table(capture.output(summary(dataframe[,variable_chr])))
             
        print(ggplot(dataframe, aes(dataframe[,variable_chr])) + geom_bar() + ggtitle(paste("Barplot of", variable_chr)) + labs(x=(variable_chr)))
        
        print(map_plot("SUBCLASS1_FARMED", "Soil Farming",dataframe, variable_chr))
         
        print(map_plot("SUBCLASS2_MULCHING", "Soil mulching",dataframe, variable_chr))
         
        print(map_plot("SUBCLASS3_IRRIGATION", "Soil Irrigation",dataframe, variable_chr))
        
        dev.off()
        return()
     }
     return()
}
```

```{r}
#make_subclass_coastal will include a new variable in the SoilData dataframe related to the coordinates of each observation. The variable is a factor which approximately classifies each observation into one of two categories: "coastal" and "interior" deppending on where the sample was extracted in relation to its proximity to the coast. For that, the first and last quartiles of the variables XUTM and YUTM were extracted and combined to generate the category "coastal", and all the leftover observations were included in "interior".  

make_subclass_coastal <- function(){

  class <- as.vector(SoilData$CLASS)
  
  coastalX <- as.vector(cut(SoilData$XUTM, breaks=c(-Inf,630679,646304,Inf), labels=c("coastal","interior","coastal")))
  
  coastalY <- as.vector(cut(SoilData$YUTM, breaks=c(-Inf,3207726,3221146,Inf), labels=c("coastal","interior","coastal")))
  
  a <- c()
  
  for (i in seq_along(coastalX)){
    
        if(class[i]=="coastal soil nonfarmed"){
          a[i] <- "coastal"
          }else if(coastalX[i]=="coastal"){
            a[i] <- "coastal"
            }else if(coastalY[i]=="coastal"){
              a[i] <- "coastal"
              }else{
                a[i] <- "interior"
              }
  }
  
  SoilData$subclass_coastal <<- as.factor(a)
  return()
  }


  
```


