---
title: "Lanzarote Soil - Functions"
author: "Yago Sánchez Gancedo"
date:
output: html_document
---

## low_cut_break and high_cut_break

A function that gives the cut points to separate data in three parts of the same size. It's intended to be used along with the "cut" function.

```{r}
low_cut_break <- function(dataframe, variable_chr){
  
  low_cut <- min(dataframe[,variable_chr], na.rm=TRUE) + ((max(dataframe[,variable_chr], na.rm=TRUE)-min(dataframe[,variable_chr], na.rm=TRUE))/3)
  
  return(low_cut)
}

high_cut_break <- function(dataframe, variable_chr){
  
  high_cut <- max(dataframe[,variable_chr], na.rm=TRUE) - ((max(dataframe[,variable_chr], na.rm=TRUE)-min(dataframe[,variable_chr], na.rm=TRUE))/3)
  
  return(high_cut)
}
```


## map_plot

Generates a plot considering the geographic distribution of a variable and a subclass clasification. The format of the data points is different deppending on the nature of the variable.

```{r}
map_plot <- function(subclass_chr, subclass_label, dataframe, variable_chr){
                
                require(ggplot2)
                require(dplyr)
   
                filtered_data <- dplyr::filter(dataframe, !is.na(dataframe[,variable_chr]))
                
                if (is.numeric(dataframe[,variable_chr])){
                        
                    ggplot(filtered_data, aes(x=XUTM, y=YUTM, size=filtered_data[,variable_chr], col = filtered_data[,subclass_chr]))+
                  geom_point()+
                  ggtitle(paste(variable_chr, "according to location and", subclass_label))+
                  labs(size=variable_chr, col=subclass_label)+
                  theme(legend.title=element_text(size=9),
                        legend.text=element_text(size=7),
                        legend.position = c(0.85,0.2),
                        legend.box="horizontal",
                        legend.key.size=unit(0.05, "cm"))  
                        
                }else if(is.factor(dataframe[,variable_chr])) {
                  ggplot(filtered_data, aes(x=XUTM, y=YUTM, shape=filtered_data[,variable_chr], col = filtered_data[,subclass_chr]))+
                    geom_point()+
                    ggtitle(paste(variable_chr, "according to location and", subclass_label))+
                    labs(shape=variable_chr, col=subclass_label)+
                    theme(legend.title=element_text(size=9),
                          legend.text=element_text(size=7),
                          legend.position = c(0.85,0.2),
                          legend.box="horizontal",
                          legend.key.size=unit(0.05, "cm"))
                }
                 
}
```

## soil_boxplot

```{r}

soil_test_boxplot <- function(subclass_chr, subclass_label, dataframe, variable_chr, comparisons){
  
  require(ggplot2)
  require(dplyr)
  
  filtered_data <- dplyr::filter(dataframe, !is.na(dataframe[,variable_chr]))
  
  if (shapiro.test(dataframe[,variable_chr])[[2]]>0.05){
    
    ggplot(filtered_data, aes(x=filtered_data[,subclass_chr], y=filtered_data[,variable_chr]))+
    geom_boxplot()+
    ggtitle(paste(variable_chr, "to", subclass_label, "(test method = t-test)" ))+
    labs(x=subclass_label, y=variable_chr)+
    stat_compare_means(comparisons=comparisons, method="t.test")

  }else{
    ggplot(filtered_data, aes(x=filtered_data[,subclass_chr], y=filtered_data[,variable_chr]))+
    geom_boxplot()+
    ggtitle(paste(variable_chr, "to", subclass_label))+
    labs(x=subclass_label, y=variable_chr)
  }
  
  
}



```

## variable_exploration

Given a variable, the function will generate a pdf with several outputs including statistics, plots, etc. with the aim of making a general exploration of the variable. It also takes into account the nature of the variable.


```{r}
# This function has still to be improved. The output is not optimal with some variables, and more exploration commands can be implemented in it.

variable_exploration <- function (dataframe, variable_chr) {
        
        require(ggplot2)
        require(dplyr)
        require(gridExtra)
        require(ggpubr)
        
        # the name of the resulting pdf file.
         filename <- paste("Exploration of ", variable_chr,".pdf", sep="") 
        
        # if the variable is numeric, a summary, histogram, test for normality and several "map plots" will be obtained.
     if(is.numeric(dataframe[,variable_chr])){
             
         pdf(filename)
         
         grid.table(capture.output(summary(dataframe[,variable_chr])))
         
         print(ggplot(dataframe, aes(dataframe[,variable_chr])) + geom_histogram() + ggtitle(paste("Histogram of", variable_chr)) + labs(x=(variable_chr)))
         
         plot.new()
         
         grid.table(capture.output(shapiro.test(dataframe[,variable_chr])))
         # MAP-PLOTS
         #map_plot("CLASS", "Soil Class", dataframe, variable_chr)??
         
         print(map_plot("SUBCLASS1_AGRICULTURALSYSTEM", "Agricultural System", dataframe, variable_chr))
         
         print(map_plot("SUBCLASS2_FARMED", "Soil Farming", dataframe, variable_chr))
         
         print(map_plot("SUBCLASS3_MULCHING", "Soil Mulching",dataframe, variable_chr))
         
         print(map_plot("SUBCLASS4_IRRIGATION", "Soil Irrigation",dataframe, variable_chr))
         
         #BOXPLOTS
         
           
           mycomparisons <- list(c("Non-Farmed", "Terraced"),c("Non-Farmed", "Tephra mulched"), c("Non-Farmed", "Sand mulched"), c("Non-Farmed", "Runoff capture"), c("Terraced", "Tephra mulched"), c("Terraced", "Sand mulched"), c("Terraced", "Runoff capture"), c("Tephra mulched", "Sand mulched"), c("Tephra mulched", "Runoff capture"), c("Sand mulched", "Runoff capture"))
           
           print(soil_test_boxplot("SUBCLASS1_AGRICULTURALSYSTEM", "Agricultural System", dataframe, variable_chr, mycomparisons))
         
           mycomparisons <- list(c("Natural", "Farmed"))
           
         print(soil_test_boxplot("SUBCLASS2_FARMED", "Soil Farming", dataframe, variable_chr, mycomparisons))
         
          mycomparisons <- list(c("No mulch", "Tephra"), c("No mulch", "Sand"), c("Tephra", "Sand"))
         
         print(soil_test_boxplot("SUBCLASS3_MULCHING", "Soil Mulching",dataframe, variable_chr, mycomparisons))
         
         mycomparisons <- list(c("No Irrigation", "Unknown"), c("No Irrigation", "Rainfed"), c("No Irrigation", "Irrigated"), c("No Irrigation", "Runoff"), c("Unknown", "Rainfed"), c("Unknown", "Irrigated"), c("Unknown", "Runoff"), c("Rainfed", "Irrigated"), c("Rainfed", "Runoff"), c("Irrigated", "Runoff"))
         
         print(soil_test_boxplot("SUBCLASS4_IRRIGATION", "Soil Irrigation",dataframe, variable_chr, mycomparisons))
           
         
         #print(soil_boxplot("SUBCLASS1_AGRICULTURALSYSTEM", "Agricultural System", dataframe, variable_chr))
         
         #print(soil_boxplot("SUBCLASS2_FARMED", "Soil Farming", dataframe, variable_chr))
         
         #print(soil_boxplot("SUBCLASS3_MULCHING", "Soil Mulching",dataframe, variable_chr))
         
         #print(soil_boxplot("SUBCLASS4_IRRIGATION", "Soil Irrigation",dataframe, variable_chr))
         
         dev.off()
         return()
     }
        
         #On the other hand, if the variable is a factor, a summary, bar plot and several "map plots" will be obtained.
         
     else if(is.factor(dataframe[,variable_chr])){
             
        pdf(filename)
             
        grid.table(capture.output(summary(dataframe[,variable_chr])))
             
        print(ggplot(dataframe, aes(dataframe[,variable_chr])) + geom_bar() + ggtitle(paste("Barplot of", variable_chr)) + labs(x=(variable_chr)))
        
        print(map_plot("SUBCLASS1_AGRICULTURALSYSTEM", "Agricultural System",dataframe, variable_chr))
        
        print(map_plot("SUBCLASS2_FARMED", "Soil Farming",dataframe, variable_chr))
         
        print(map_plot("SUBCLASS3_MULCHING", "Soil mulching",dataframe, variable_chr))
         
        print(map_plot("SUBCLASS4_IRRIGATION", "Soil Irrigation",dataframe, variable_chr))
        
        dev.off()
        return()
     }
     return()
}
```

```{r}
#make_subclass_coastal will include a new variable in the SoilData dataframe related to the coordinates of each observation. The variable is a factor which approximately classifies each observation into one of two categories: "coastal" and "interior" deppending on where the sample was extracted in relation to its proximity to the coast. For that, the first and last quartiles of the variables XUTM and YUTM were extracted and combined to generate the category "coastal", and all the leftover observations were included in "interior". This is just an approximation an it is subject to improving.

make_subclass_coastal <- function(){
  
  class <- as.vector(SoilData$CLASS)
  
  coastalX <- as.vector(cut(SoilData$XUTM, breaks=c(-Inf,630679,646304,Inf), labels=c("coastal","interior","coastal")))
  
  coastalY <- as.vector(cut(SoilData$YUTM, breaks=c(-Inf,3207726,3221146,Inf), labels=c("coastal","interior","coastal")))
  
  a <- c()
  
  for (i in seq_along(coastalX)){
    
        if(class[i]=="coastal soil nonfarmed"){
          a[i] <- "coastal"
          }else if(coastalX[i]=="coastal"){
            a[i] <- "coastal"
            }else if(coastalY[i]=="coastal"){
              a[i] <- "coastal"
              }else{
                a[i] <- "interior"
              }
  }
  
  SoilData$subclass_coastal <<- as.factor(a)
  return()
  }


  
```

## Remove NA lines

This function removes all the lines of a dataframe that contain any NA values for the columns passed as argument.

```{r}
Remove_NA_lines <- function(dataframe, columns_num){
  eliminated_rows <- c()
  for (i in columns){
    for (a in seq_along(dataframe[,i])){
      if (is.na(dataframe[,i][a])){
        if (a %in% eliminated_rows) next
        else eliminated_rows <- c(eliminated_rows, a)
        }
      }
  }
  new_dataframe <- dataframe[-eliminated_rows,]
  return(new_dataframe)
  }

```

## Remove outliers

This function takes as an outlier every observation whose standard deviation is greater than 6, then removes that observation from the dataframe.

```{r}
Remove_outliers <- function(dataframe, columns_num){
  
      for (i in columns_num){
        Q1 <- quantile(dataframe[,i], probs=0.25, na.rm=TRUE)
        Q3 <- quantile(dataframe[,i], probs=0.75, na.rm=TRUE)
        low_value <- Q1 - IQR(dataframe[,i], na.rm=TRUE)*1.5
        high_value <- Q3 + IQR(dataframe[,i], na.rm=TRUE)*1.5
        
        for (a in seq_along(dataframe[,i])){
          if (is.na(dataframe[,i][a])) next
          else if (dataframe[,i][a] < low_value) dataframe[,i][a]<- NA
          else if (dataframe[,i][a] > high_value) dataframe[,i][a]<- NA
        }
      }
      return(dataframe)
  }
````


