---
title: "Lanzarote Soil"
author: "Yago Sánchez Gancedo"
date:
output: html_document
---

```{r}
#Path to access the local folder where the script with the functions associated with this work is

folderScripts <- "/Users/cacharra2.0/Desktop/TFM/R/LanzaroteSoil/"
source(file=file.path(folderScripts, "Lanzarote Soil - Functions.Rmd"))

```


## 1. Introduction of the data

The first task is to read the data in R

```{r}
install.packages("openxlsx", repos = "http://cran.us.r-project.org")
library(openxlsx)

#Creating the local path to the data, and using that path to generate an object containing such data

DataPath <- "/Users/cacharra2.0/Desktop/TFM/Lanzarotesoil.xlsx"
SoilData <- read.xlsx(DataPath, colNames = TRUE)
```

## 2. Superficial exploration of the data

Once the data is stored in the variable "SoilData", a brief superficial exploration can be performed to ensure that everything is correct and that the data is treated accordingly:

Visualization of the first and last subset of data, to check if it matches with the original data table.
```{r}
head(SoilData)
tail(SoilData)
```

Structure of the data.

```{r}
dim(SoilData)
ncol(SoilData)
nrow(SoilData)
str(SoilData)
```

A summary to see if the variables are treated with the correct format

```{r}
summary(SoilData)
```

As it can be seen with the two previous commands, some variables are not treated as desired. A correction is needed:

```{r}
SoilData$SAMPLE.NUMBER <- as.character(SoilData$SAMPLE.NUMBER)
SoilData$CLASS <- factor(SoilData$CLASS)
SoilData$SUBCLASS1_AGRICULTURALSYSTEM <- factor(SoilData$SUBCLASS1_AGRICULTURALSYSTEM )
SoilData$SUBCLASS2_FARMED <- factor(SoilData$SUBCLASS2_FARMED)
SoilData$SUBCLASS3_MULCHING <- factor(SoilData$SUBCLASS3_MULCHING)
SoilData$SUBCLASS4_IRRIGATION <- factor(SoilData$SUBCLASS4_IRRIGATION)
SoilData$CLASS.REFERENCE <- factor(SoilData$CLASS.REFERENCE)
SoilData$DEPTH <- factor(SoilData$DEPTH) #LABELS!
SoilData$TEXTURE <- factor(SoilData$TEXTURE) #LABELS!
```

## 3. Wrangling of the data

Creating a new factor with the different pH types, and categories for rainfall, slope and elevation.

```{r}
SoilData$PHgrouped <- cut(SoilData$PH, breaks=c(-Inf, 6, 8, Inf), labels=c("Acid", "Neutral", "Basic"))

SoilData$rainfallgrouped <- factor(cut(SoilData$RAINFALL_A, breaks=c(-Inf,low_cut_break(SoilData,"RAINFALL_A",3), high_cut_break(SoilData,"RAINFALL_A",3), Inf), labels=c("local low rainfall", "local medium rainfall", "local high rainfall")))

SoilData$slopegrouped <- factor(cut(SoilData$SLOPE, breaks=c(-Inf,5,15, Inf), labels=c("low slope", "medium slope", "high slope")))

SoilData$elevatgrouped <- factor(cut(SoilData$ELEVAT, breaks=c(-Inf,low_cut_break(SoilData,"ELEVAT",3), high_cut_break(SoilData,"ELEVAT",3), Inf), labels=c("low elevation", "normal elevation", "high elevation")))
```



## 4. Variable exploration

The function "variable_exploration" is applied to all variables, which generated a pdf file with all the information required.

```{r}
variable_names <- colnames(SoilData)
for (i in seq_along(variable_names)){
  variable_exploration(SoilData, variable_names[i])
}
```
