---
title: "Lanzarote Soil"
author: "Yago Sánchez Gancedo"
date:
output: html_document
---

```{r, include=FALSE}
#Path to access the local folder where the script with the functions associated with this work is

#folderScripts <- "/Users/cacharra2.0/Desktop/TFM/R/LanzaroteSoil/"
#source(file=file.path(folderScripts, "LanzaroteSoil.Rproj"))
library(rmarkdown)
render("Lanzarote Soil - Functions.Rmd")
```


## 1. Introduction of the data

The first task is to read the data in R

```{r}
install.packages("openxlsx", repos = "http://cran.us.r-project.org")
library(openxlsx)

#Creating the local path to the data, and using that path to generate an object containing such data

DataPath <- "/Users/cacharra2.0/Desktop/TFM/Lanzarotesoil.xlsx"
 SoilData <- read.xlsx(DataPath, colNames = TRUE)
```

## 2. Superficial exploration of the data

Once the data is stored in the variable "SoilData", a brief superficial exploration can be performed to ensure that everything is correct and that the data is treated accordingly:

Visualization of the first and last subset of data, to check if it matches with the original data table.
```{r}
head(SoilData)
tail(SoilData)
```

Structure of the data.

```{r}
dim(SoilData)
ncol(SoilData)
nrow(SoilData)
str(SoilData)
```

A summary to see if the variables are treated with the correct format

```{r}
summary(SoilData)
```

As it can be seen with the two previous commands, some variables are not treated as desired. A correction is needed:

```{r}
SoilData$SAMPLE.NUMBER <- as.character(SoilData$SAMPLE.NUMBER)
SoilData$CLASS <- factor(SoilData$CLASS)
SoilData$SUBCLASS1_AGRICULTURALSYSTEM <- factor(SoilData$SUBCLASS1_AGRICULTURALSYSTEM )
SoilData$SUBCLASS2_FARMED <- factor(SoilData$SUBCLASS2_FARMED)
SoilData$SUBCLASS3_MULCHING <- factor(SoilData$SUBCLASS3_MULCHING)
SoilData$SUBCLASS4_IRRIGATION <- factor(SoilData$SUBCLASS4_IRRIGATION)
SoilData$CLASS.REFERENCE <- factor(SoilData$CLASS.REFERENCE)
SoilData$DEPTH <- factor(SoilData$DEPTH, levels=c(1,2,3,4), labels=c("0-30 cm", "30-60 cm", "60-90 cm", ">90 cm"))
SoilData$TEXTURE <- factor(SoilData$TEXTURE, levels= c(1,2,3,4), labels=c("Arenosa", "Limosa", "Arcillosa", "Franca"))
```



## 3. Wrangling of the data

Creating a new factor with the different pH types, and categories for rainfall, slope and elevation.

```{r}
SoilData$PHgrouped <- cut(SoilData$PH, breaks=c(-Inf, 6, 8, Inf), labels=c("Acid", "Neutral", "Basic"))

SoilData$rainfallgrouped <- factor(cut(SoilData$RAINFALL_A, breaks=c(-Inf,low_cut_break(SoilData,"RAINFALL_A"), high_cut_break(SoilData,"RAINFALL_A"), Inf), labels=c("local low rainfall", "local medium rainfall", "local high rainfall")))

SoilData$slopegrouped <- factor(cut(SoilData$SLOPE, breaks=c(-Inf,5,15, Inf), labels=c("low slope", "medium slope", "high slope")))

SoilData$elevatgrouped <- factor(cut(SoilData$ELEVAT, breaks=c(-Inf,low_cut_break(SoilData,"ELEVAT"), high_cut_break(SoilData,"ELEVAT"), Inf), labels=c("low elevation", "normal elevation", "high elevation")))
```


## 4. Variable exploration

The function "variable_exploration" is applied to all variables, which generated a pdf file with all the information required.

```{r, echo=T, results="hide"}
variable_names <- colnames(SoilData)
for (i in seq_along(variable_names)){
  variable_exploration(SoilData, variable_names[i])
}
```

In the different pdf files that variable_exploration generated, some outliers can be seen in the boxplots. These are removed below manually via the remove_outliers_visually function, so they don't interfere in the next steps of the analysis.

```{r}
SoilData <- Remove_outliers_visually(SoilData, "Caexch", 3, 0)
SoilData <-Remove_outliers_visually(SoilData, "CEC", 3, 0)
SoilData <-Remove_outliers_visually(SoilData, "Corg", 2, 0)
SoilData <-Remove_outliers_visually(SoilData, "CU", 2, 0)
SoilData <-Remove_outliers_visually(SoilData, "EC", 4, 0)
SoilData <-Remove_outliers_visually(SoilData, "FE", 1, 0)
SoilData <-Remove_outliers_visually(SoilData, "HWSB", 3, 0)
SoilData <-Remove_outliers_visually(SoilData, "Kexch", 1, 0)
SoilData <-Remove_outliers_visually(SoilData, "Mgexch", 1, 0)
SoilData <-Remove_outliers_visually(SoilData, "MN", 6, 0)
SoilData <-Remove_outliers_visually(SoilData, "Naexch", 4, 0)
SoilData <-Remove_outliers_visually(SoilData, "PH", 0, 3)
SoilData <-Remove_outliers_visually(SoilData, "PHwater", 0, 5)
SoilData <-Remove_outliers_visually(SoilData, "Polsen", 4, 0)
SoilData <-Remove_outliers_visually(SoilData, "SAR", 3, 0)
SoilData <-Remove_outliers_visually(SoilData, "Silt.g.kg-1", 1, 0)
SoilData <-Remove_outliers_visually(SoilData, "TN", 4, 0)
SoilData <-Remove_outliers_visually(SoilData, "ZN", 6, 0)
```


### CONCLUSIONS FROM VARIABLE EXPLORATION




## 5. Principal component analysis

```{r}
##REMOVE ALL THIS CHUNK???
numeric_variables <- c()
for (i in seq_along(variable_names)){
  
  variable_name <- variable_names[i]
  
  if (is.numeric(SoilData[,variable_name])){
    numeric_variables <- c(numeric_variables, variable_name)
  }
}

numeric_SoilData <- dplyr::select(SoilData, (numeric_variables))

#Coordinates variables are also extracted

numeric_SoilData$XUTM <- NULL
numeric_SoilData$YUTM <- NULL
```

The first principal component analysis will be performed only with columns 13 to 32, as those are all numeric, descriptive of soil quality and have very few NA's in comparison with other variables.

```{r}
SoilData_pca1filter <- Remove_NA_lines(SoilData, c(14:33))

SoilData.pca1 <- prcomp(SoilData_pca1filter[,14:33], center=TRUE, scale.=TRUE)
library(ggbiplot)
```


```{r}
ggbiplot(SoilData.pca1, ellipse=TRUE, groups = SoilData_pca1filter$SUBCLASS1_AGRICULTURALSYSTEM)
```

```{r}
ggbiplot(SoilData.pca1, ellipse=TRUE, groups = SoilData_pca1filter$SUBCLASS2_FARMED)
```


```{r}
ggbiplot(SoilData.pca1, ellipse=TRUE, groups = SoilData_pca1filter$SUBCLASS3_MULCHING)
```

```{r}
ggbiplot(SoilData.pca1, ellipse=TRUE, groups = SoilData_pca1filter$SUBCLASS4_IRRIGATION)
```

For the PCA grouping by position of the sample (inland, mountain or coastal), the variables PHwater and HWSB (Hot Water Soluble Boro) will not be taken into account as there are too many NAs for the coastal category.

```{r}
SoilData_pca1filter2 <- Remove_NA_lines(SoilData, c(14,15,17:31,33))

SoilData.pca1position <- prcomp(SoilData_pca1filter2[,c(14,15,17:31,33)], center=TRUE, scale.=TRUE)

ggbiplot(SoilData.pca1position, ellipse=TRUE, groups = SoilData_pca1filter2$SUBCLASS5_POSITION)
```

Below, the PCA is repeated, taking into account all the variables

```{r}
SoilData_pca1filterall <- Remove_NA_lines(SoilData, c(14:33,37:39))

SoilData.pca2 <- prcomp(SoilData_pca1filterall[,c(14:33,37:39)], center=TRUE, scale.=TRUE)
library(ggbiplot)

```

```{r}
ggbiplot(SoilData.pca2, ellipse=TRUE, groups = SoilData_pca1filterall$SUBCLASS1_AGRICULTURALSYSTEM)
```

```{r}
ggbiplot(SoilData.pca2, ellipse=TRUE, groups = SoilData_pca1filterall$SUBCLASS2_FARMED)
```


```{r}
ggbiplot(SoilData.pca2, ellipse=TRUE, groups = SoilData_pca1filterall$SUBCLASS3_MULCHING)
```

```{r}
ggbiplot(SoilData.pca2, ellipse=TRUE, groups = SoilData_pca1filterall$SUBCLASS4_IRRIGATION)
```
```{r}
SoilData_pca1filterall2 <- Remove_NA_lines(SoilData, c(14,15,17:31,33,37:39))

SoilData.pca1position2 <- prcomp(SoilData_pca1filterall2[,c(14,15,17:31,33,37:39)], center=TRUE, scale.=TRUE)

ggbiplot(SoilData.pca1position2, ellipse=TRUE, groups = SoilData_pca1filterall2$SUBCLASS5_POSITION)
```