---
title: "Lanzarote Soil - Functions"
author: "Yago Sánchez Gancedo"
date:
output: html_document
---


## map_plot

Generates a plot considering the geographic distribution of a variable and a subclass clasification. The format of the data points is different deppending on the nature of the variable.

```{r}
map_plot <- function(subclass_chr, subclass_label, dataframe, variable_chr){
                
                require(ggplot2)
                require(dplyr)
   
                filtered_data <- dplyr::filter(dataframe, !is.na(dataframe[,variable_chr]))
                
                if (is.numeric(dataframe[,variable_chr])){
                        
                    print(ggplot(filtered_data, aes(x=XUTM, y=YUTM, size=filtered_data[,variable_chr], col = filtered_data[,subclass_chr])) 
                       + geom_point() 
                       + ggtitle(paste(variable_chr, "according to location and", subclass_label)) 
                       + labs(size=variable_chr, col=subclass_label)
                       + theme(legend.title=element_text(size=9),
                               legend.text=element_text(size=7),
                               legend.position = c(0.85,0.2),
                               legend.box="horizontal",
                               legend.key.size=unit(0.05, "cm")))  
                        
                }else if(is.factor(dataframe[,variable_chr])) {
                 
                print(ggplot(filtered_data, aes(x=XUTM, y=YUTM, shape=filtered_data[,variable_chr], col = filtered_data[,subclass_chr])) 
                       + geom_point() 
                       + ggtitle(paste(variable_chr, "according to location and", subclass_label)) 
                       + labs(shape=variable_chr, col=subclass_label)
                       + theme(legend.title=element_text(size=9),
                               legend.text=element_text(size=7),
                               legend.position = c(0.85,0.2),
                               legend.box="horizontal",
                               legend.key.size=unit(0.05, "cm")))
                }
                 
}
```


## variable_exploration

Given a variable, the function will generate a pdf with several outputs including statistics, plots, etc. with the aim of making a general exploration of the variable. It also takes into account the nature of the variable.


```{r}
variable_exploration <- function (dataframe, variable_chr) {
        
        require(ggplot2)
        require(dplyr)
        require(gridExtra)
        
        # the name of the resulting pdf file.
         filename <- paste("Exploration of ", variable_chr,".pdf", sep="") 
        
        # if the variable is numeric, a summary, histogram, test for normality and several "map plots" will be obtained.
     if(is.numeric(dataframe[,variable_chr])){
             
         pdf(filename)
         
         grid.table(capture.output(summary(dataframe[,variable_chr])))
         
         print(ggplot(dataframe, aes(dataframe[,variable_chr])) + geom_histogram() + ggtitle(paste("Histogram of", variable_chr)) + labs(x=(variable_chr)))
         
         plot.new()
         
         grid.table(capture.output(shapiro.test(dataframe[,variable_chr])))
         
         #map_plot("CLASS", "Soil Class", dataframe, variable_chr)??
         
         map_plot("SUBCLASS1_FARMED", "Soil Farming", dataframe, variable_chr)
         
         map_plot("SUBCLASS2_MULCHING", "Soil mulching",dataframe, variable_chr)
         
         map_plot("SUBCLASS3_IRRIGATION", "Soil Irrigation",dataframe, variable_chr)
         
         dev.off()
         return()
     }
        
         #On the other hand, if the variable is a factor, a summary, bar plot and several "map plots" will be obtained.
         
     else if(is.factor(dataframe[,variable_chr])){
             
        pdf(filename)
             
        grid.table(capture.output(summary(dataframe[,variable_chr])))
             
        print(ggplot(dataframe, aes(dataframe[,variable_chr])) + geom_bar() + ggtitle(paste("Barplot of", variable_chr)) + labs(x=(variable_chr)))
        
        map_plot("SUBCLASS1_FARMED", "Soil Farming",dataframe, variable_chr)
         
        map_plot("SUBCLASS2_MULCHING", "Soil mulching",dataframe, variable_chr)
         
        map_plot("SUBCLASS3_IRRIGATION", "Soil Irrigation",dataframe, variable_chr)
        
        dev.off()
        return()
     }
     return()
}
```

 
